---
title: "oWell-3"
output: html_document
---

---
title: "o-well #3"
author: "Quynh Tran  [quynh_tran]"
output: html_document
---

## Reading the saudi.txt file that contains all the data of 23 wells obtained from o-well assignment 1, including fault, geology, metal types, chemical parameters.
```{r}
library(readr)
saudi <- read_delim("https://raw.githubusercontent.com/quynh-tran419/WSU_STAT419/master/workspace/o-well/dataset/saudi.txt", "|", escape_double = FALSE )
saudi = data.frame(saudi)
saudi
metals <- read_delim("https://raw.githubusercontent.com/quynh-tran419/WSU_STAT419/master/workspace/o-well/dataset/metals.txt", "|", escape_double = FALSE )

```


## It is mentioned that northern located wells (Th mean is 1549mg/L) tend to have a higher level of hardness compared to southern located wells (TH mean is 1098mg/L). So how does the amount of Calcium cations relate to where the wells located and theirs hardness?
## The bubble chart shows that the more Calcium cations present in the water, the higher the level of water hardness is.


```{r, fig.width=9, fig.height=7.5}

library(ggplot2)
library(dplyr)
ggplot(saudi, aes(latitude,TH,size = Ca, color = factor(well))) +
  geom_point(alpha =0.7) +
  scale_size(range = c(0.1,20), name = "Ca2+") +
  ggtitle("The correlation between the latitude,total hardness,and the amount of dissolved Ca2+ of each well") +
  xlab("Latitude") + ylab("Total Hardness") 

```

# With the statistics for dissolved metals results of groundwater in 23 wells, how do I generally indicate the amount of each metal present in each well? 
# This connected scatter chart generally shows how much each metal presents and the major metal presents in each well. For example, well #1 contains the most Fe while well #23 contains the least. However, since there are many data overlapping, this plot does not precisely tell me the maximum or minimum amount of each metal in the water sample.

``````{r, fig.width=9, fig.height=7.5}

plot( saudi$well, saudi$As , type="b" , bty="l" , xlab="well" , ylab="μg/L" , col=rgb(0.2,0.4,0.1,0.7) , lwd=3 , pch=17 , ylim=c(1,100) )
lines(saudi$well,saudi$B , col=rgb(0.8,0.4,0.1,0.7) , lwd=3 , pch=19 , type="b" )
lines(saudi$well,saudi$Ba , col=rgb(0.5,0.4,0.1,0.7) , lwd=3 , pch=15 , type="b" )
lines(saudi$well,saudi$Be , col=rgb(0.9,0.4,0.2,0.7) , lwd=3 , pch=10 , type="b" )
lines(saudi$well,saudi$Cd , col=rgb(0.7,0.4,0.4,0.7) , lwd=3 , pch=9 , type="b" )
lines(saudi$well,saudi$Co , col=rgb(0.6,0.7,0.1,0.7) , lwd=3 , pch=11 , type="b" )
lines(saudi$well,saudi$Cr , col=rgb(0.5,0.4,0.1,0.7) , lwd=3 , pch=12 , type="b" )
lines(saudi$well,saudi$cu , col=rgb(0.4,0.9,0.1,0.7) , lwd=3 , pch=13 , type="b" )
lines(saudi$well,saudi$Fe , col=rgb(0.3,0.4,0.1,0.7) , lwd=3 , pch=14 , type="b" )
lines(saudi$well,saudi$Hg , col=rgb(0.2,0.4,0.1,0.7) , lwd=3 , pch=7 , type="b" )
lines(saudi$well,saudi$Mn , col=rgb(0.1,0.4,0.1,0.6) , lwd=3 , pch=8 , type="b" )
lines(saudi$well,saudi$Mo , col=rgb(1.0,0.4,0.1,0.7) , lwd=3 , pch=6 , type="b" )
lines(saudi$well,saudi$Pb , col=rgb(0.8,0.6,0.5,0.4) , lwd=3 , pch=5 , type="b" )
lines(saudi$well,saudi$Se , col=rgb(0.8,0.8,0.8,0.8) , lwd=3 , pch=4 , type="b" )
lines(saudi$well,saudi$Zn , col=rgb(0.8,0.4,0.9,0.7) , lwd=3 , pch=3 , type="b" )
legend("topleft",
legend= c("As","B","Ba","Be","Cd","Co","Cr","Cu","Fe","Hg","Mn","Mo","Pb","Se","Zn"),
col = c(rgb(0.2,0.4,0.1,0.7), 
  rgb(0.8,0.4,0.1,0.7), rgb(0.9,0.4,0.2,0.7),col=rgb(0.7,0.4,0.4,0.7),col=rgb(0.6,0.7,0.1,0.7),col=rgb(0.5,0.4,0.1,0.7),col=rgb(0.4,0.9,0.1,0.7),col=rgb(0.3,0.4,0.1,0.7),col=rgb(0.2,0.4,0.1,0.7),col=rgb(0.1,0.4,0.1,0.6),col=rgb(1.0,0.4,0.1,0.7),col=rgb(0.8,0.6,0.5,0.4),col=rgb(0.8,0.8,0.8,0.8),col=rgb(0.8,0.4,0.9,0.7)
  ), title= "Dissolved metals vs. Well",
  pch = c(17,19,15,10,9,11,12,13,14,7,8,6,5,4,3), 
  bty = "n", 
  pt.cex = 1, 
  cex = 0.7, 
  text.col = "black", 
  horiz = T , 
  inset = c(0.1, 0.1,0.1,0.1, 0.1,0.1,0.1, 0.1,0.1,0.1, 0.1,0.1,0.1, 0.1))


```
# Regarding the maximum and the minimum amount of each metal in each well, how much does each metal exceed from the average? To have a better insights of the maximum, minimum, and the average amount of each metal presenting in the water sample, I made this scatter plot to have a better visualization of those variables. As presented, the maximum amount of Ba is really high compared to the average amount of it. This leads to another question, what makes the exceeding amount so big? 

```{r}
library(pastecs)
library(reshape2)
library(tidyverse)
metals = owell.metals[c(1:23),-c(16:21)]

metals = as.data.frame(sapply(metals, as.numeric))
mvs <- stat.desc(metals)
mvs
round(mvs, 2)
mvs_max = mvs[5,]
mvs_min = mvs[4,]
mvs_mean = mvs[9,]
mvs_min <- melt(mvs_min)
mvs_max <- melt(mvs_max)
mvs_mean <- melt(mvs_mean)
mvs_max
mvs_min
ggplot() +
  geom_point(data= mvs_max , aes(x= variable, y=value),shape = 20, color = "blue", size = 3) + 
  geom_point(data =mvs_min, aes(x= variable, y= value), color ="red", size =2, alpha = 0.7)+ 
  geom_point(data =mvs_mean, aes(x= variable, y= value), color ="green", size =2, alpha = 0.7)+ 
  ggtitle("The maximum amount of each metal presents in the water sample") +
  xlab("Type of metal") +
  ylab("μg/L")

```


## How is the distance from each well related to its pH? 
## The circular barplot is created to desmonstrate how the variables might be corresponded. As seen below, the pH of the 23 wells are not diversely distributed. The lowest pH is 7 and the highest pH is only 7.8, which indicates the water sample from the wells are suitable for drinking in terms of the pH level. Also, according to the plot, wells that are closer to a fault tend to have higher pH levels (blue-ish bins) while the further wells from a fault have lower pH (yellow-ish well with pH 7.2).
```{r}
library(viridis)
library(tidyverse)
owell.chemistry$pH
pH <- data_frame(owell.chemistry$well, owell.chemistry$pH )
ggplot(owell.chemistry, aes(x= well, y = pH, fill = fault)) +
  geom_bar(stat = "identity", alpha = 1)+
  scale_fill_viridis(option = "C")+
  ylim(-3,8) +
  coord_polar(start = 0)+ 
  geom_text(data = owell.chemistry, aes(label = pH, hjust = "hjust"), color= "red")
  


```

## What is the correlation between altitude and rock formations?
## This chart is meant to show the relationship between altitude and rock formations. The most popular rock forms are sand and gypsum while the least are alkaline-granite and volcanic rock-granite.
```{r}

geo = data_frame(geology, owell.chemistry$altitude.ft)
geo
ggplot(geo, aes(y= geology, x= owell.chemistry$altitude.ft), fill = "sky blue") +
  geom_point(size=3, alpha =0.7)+
  scale_x_continuous(n.breaks = 20)+
  ggtitle("Geology vs. Altitude") +
  xlab("Altitude(ft)")

```